Measurement Bias from Address Aliasing

### 2014-05-30 paper.tex
Fix spellcheck errors

\section{Introduction}
Accurately measuring the performance of computer programs is important in order to evaluate new algorithms or compare different implementations.
A key challenge for performance analysist and systems researchers is handling \emph{measurement bias}.
Changes to external factors of the system, such as link ordering or the contents of environment variables, has been shown to have potentially significant impacts on performance in real applications~\cite{Mytkowicz:2009:WrongData}.
The underlying explanation is likely that this changes memory addresses of code or data, which in turn interacts with various low level hardware buffers at runtime.
Because of the complexity of modern processors, it is often hard or even impossible to predict exactly how changes to for example system environment variables ultimately affect program execution.

@@

\section{Related Work}
\label{sec:related}
Measurement bias and observer effect in performance analysis has been studied in detail by Mytkowicz et al~\shortcite{Mytkowicz:2008:OE&MB,Mytkowicz:2008:Easy}.
The authors introduce environment variables and link ordering as examples of external bias triggers, showing how standardised {\small SPECint} benchmarks suffer from measurement bias when trying to evaluate the effectiveness of O3 over O2. 
Strategies for detecting bias through causal analysis and randomization of experimental setups is introduced.
The microkernel example we analyzed is taken from a followup paper by the same authors, presenting similar results for an older Core 2 processor~\cite{Mytkowicz:2009:WrongData}.
Assuming the root causes of measurement bias were the same, our results indicate that address aliasing issues is probably relevant on several previous generations of Intel architectures as well.

>>>>

\section{Introduction}
Accurately measuring the performance of computer programs is important in order to evaluate new algorithms or compare different implementations.
A key challenge for performance analysts and systems researchers is handling \emph{measurement bias}.
Changes to external factors of the system, such as link ordering or the contents of environment variables, has been shown to have potentially significant impacts on performance in real applications~\cite{Mytkowicz:2009:WrongData}.
The underlying explanation is likely that this changes memory addresses of code or data, which in turn interacts with various low level hardware buffers at runtime.
Because of the complexity of modern processors, it is often hard or even impossible to predict exactly how changes to for example system environment variables ultimately affect program execution.

@@

\section{Related Work}
\label{sec:related}
Measurement bias and observer effect in performance analysis has been studied in detail by Mytkowicz et al~\shortcite{Mytkowicz:2008:OE&MB,Mytkowicz:2008:Easy}.
The authors introduce environment variables and link ordering as examples of external bias triggers, showing how standardized {\small SPECint} benchmarks suffer from measurement bias when trying to evaluate the effectiveness of O3 over O2. 
Strategies for detecting bias through causal analysis and randomization of experimental setups is introduced.
The microkernel example we analyzed is taken from a followup paper by the same authors, presenting similar results for an older Core 2 processor~\cite{Mytkowicz:2009:WrongData}.
Assuming the root causes of measurement bias were the same, our results indicate that address aliasing issues is probably relevant on several previous generations of Intel architectures as well.
